---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

### Data in a dataframe
```{r}

ID = c("Riccardo", "Kristian", "Josh", "Mikkel")
Correct = c(3,2, 160, 66)
Total = c(6, 2, 198, 132)


data = data.frame(ID, Correct, Total)

```

Questions:

1. What's Riccardo's estimated knowledge of CogSci? 0.5, chance level

What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]? 0.5, chance level

- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
#Define the grid
dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- rep( 1 , dens ) # Flat
# prior <- ifelse(p_grid < 0.5 , 0 , 1) #informed
# prior <- exp(-5*abs( p_grid - 0.5) #conservative

#Compute the likelihood at each value in grid
likelihood_R <- dbinom( 3 , size=6 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_R <- likelihood_R * prior

#Standardize the posterior (so it sums to 1)
posterior_R <- unstd.posterior_R/ sum(unstd.posterior_R)

#Draw the plot
grid_data_R=data.frame(grid=p_grid,posterior=posterior_R,prior=prior,likelihood=likelihood_R)

library(ggplot2)
ggplot(grid_data_R,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

#probabiltiy of knowing more than chance
sum(posterior[p_grid < 0.5])


#Quadratic approximation
library(rethinking)

quadR = map(
  alist(
    c ~ dbinom(6,p),
    p ~ dunif(0,1)
  ),
  data=list(c=3))

precise(quadR)

c = 3
n = 6
curve(dbeta(x, c+1, n-c+1), from=0, to=1)
curve(dnorm(x, 0.5, 0.2), lty=2, add=TRUE)

```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. 
Comment on and compare the posteriors of Riccardo and Mikkel.

## Because Mikkel has answered more questions than Riccardo, the graph is more narrow/focused to the probability of 0.5. A new data point would not change Mikkels graph but would have a radical influence on Riccardo's graph. 

2a. Produce plots of the prior, and posterior for each teacher.
```{r}
#Kristian

dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- rep( 1 , dens ) # Flat
# prior <- ifelse(p_grid < 0.5 , 0 , 1) #informed
# prior <- exp(-5*abs( p_grid - 0.5) #conservative

#Compute the likelihood at each value in grid
likelihood_K <- dbinom( 2 , size=2 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_K <- likelihood_K * prior

#Standardize the posterior (so it sums to 1)
posterior_K <- unstd.posterior_K / sum(unstd.posterior_K)


#Draw the plot
grid_data_K=data.frame(grid=p_grid,posterior=posterior_K,prior=prior,likelihood=likelihood_K)

ggplot(grid_data_K,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

#Josh

dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- rep( 1 , dens ) # Flat
# prior <- ifelse(p_grid < 0.5 , 0 , 1) #informed
# prior <- exp(-5*abs( p_grid - 0.5) #conservative

#Compute the likelihood at each value in grid
likelihood_J <- dbinom( 160 , size=198 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_J <- likelihood_J * prior

#Standardize the posterior (so it sums to 1)
posterior_J <- unstd.posterior_J / sum(unstd.posterior_J)


#Draw the plot
grid_data_J=data.frame(grid=p_grid,posterior=posterior_J,prior=prior,likelihood=likelihood_J)

ggplot(grid_data_J,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

#Mikkel
dens=80
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- rep( 1 , dens ) # Flat
# prior <- ifelse(p_grid < 0.5 , 0 , 1) #informed
# prior <- exp(-5*abs( p_grid - 0.5) #conservative

#Compute the likelihood at each value in grid
likelihood_M <- dbinom( 66 , size=132 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_M <- likelihood_M * prior

#Standardize the posterior (so it sums to 1)
posterior_M <- unstd.posterior_M / sum(unstd.posterior_M)

#Draw the plot
grid_data_M=data.frame(grid=p_grid,posterior=posterior_M,prior=prior,likelihood=likelihood_M)

ggplot(grid_data_M,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?

3a. Produce plots of the prior and posterior for each teacher.

```{r}
#prior change Riccardo
dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- dnorm( p_grid, mean=0.8, sd=0.2) 

#Compute the likelihood at each value in grid
likelihood_R <- dbinom( 3 , size=6 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_R <- likelihood_R * prior

#Standardize the posterior (so it sums to 1)
posterior_R <- unstd.posterior_R / sum(unstd.posterior_R)

#Draw the plot
grid_data_R=data.frame(grid=p_grid,posterior=posterior_R,prior=prior,likelihood=likelihood_R)

ggplot(grid_data_R,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")


#prior change Kristian
dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior <- dnorm( p_grid, mean=0.8, sd=0.2)

#Compute the likelihood at each value in grid
likelihood_K <- dbinom( 2 , size=2 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_K <- likelihood_K * prior

#Standardize the posterior (so it sums to 1)
posterior_K <- unstd.posterior_K / sum(unstd.posterior_K)

#Draw the plot
grid_data_K=data.frame(grid=p_grid,posterior=posterior_K,prior=prior,likelihood=likelihood_K)

ggplot(grid_data_K,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")


#prior change in Josh
dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_J <- dnorm(p_grid, mean = 0.8, sd=0.2)

#Compute the likelihood at each value in grid
likelihood_J <- dbinom( 160 , size=198 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_J <- likelihood_J * prior_J

#Standardize the posterior (so it sums to 1)
posterior_J <- unstd.posterior_J / sum(unstd.posterior_J)

#Draw the plot
grid_data_J=data.frame(grid=p_grid,posterior=posterior_J,prior=prior_J,likelihood=likelihood_J)

ggplot(grid_data_J,aes(grid,posterior))+ geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

#prior change in Mikkel
dens=80
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_M <- dnorm(p_grid, mean= 0.8, sd=0.2)

#Compute the likelihood at each value in grid
likelihood_M <- dbinom( 66 , size=132 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_M <- likelihood_M * prior_M

#Standardize the posterior (so it sums to 1)
posterior_M <- unstd.posterior_M / sum(unstd.posterior_M)

#Draw the plot
grid_data_M=data.frame(grid=p_grid,posterior=posterior_M,prior=prior_M,likelihood=likelihood_M)

ggplot(grid_data_M,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
```
Do the results change (and if so how)?
The plots show that the prior have a large influence on particpants with a small amount of questions, but does only have a small infleunce on particpants with many questions. 
The more questions the teachers has answered the less infleunce the prior has on the posterior. 
Having a gaussian distributed prior affects the distribution of the posterior in a gaussian direction. 


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
#prior change Riccardo
dens=100
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_R <- dnorm( p_grid, mean=0.8, sd=0.2) 
#prior_R <- rep( 1 , dens ) #flat

#Compute the likelihood at each value in grid
likelihood_R <- dbinom( 300 , size=600 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_R <- likelihood_R * prior_R

#Standardize the posterior (so it sums to 1)
posterior_R <- unstd.posterior_R / sum(unstd.posterior_R)

#Draw the plot
grid_data_R=data.frame(grid=p_grid,posterior=posterior_R,prior=prior_R,likelihood=likelihood_R)

ggplot(grid_data_R,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

```


```{r}
#prior change Kristian
dens=100
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_K <- dnorm( p_grid, mean=0.8, sd=0.2)
#prior_K <- rep( 1 , dens ) #flat

#Compute the likelihood at each value in grid
likelihood_K <- dbinom( 200 , size=200 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_K <- likelihood_K * prior_K

#Standardize the posterior (so it sums to 1)
posterior_K <- unstd.posterior_K / sum(unstd.posterior_K)

#Draw the plot
grid_data_K=data.frame(grid=p_grid,posterior=posterior_K,prior=prior_K,likelihood=likelihood_K)

ggplot(grid_data_K,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
```


```{r}
#prior change in Josh
dens=20
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_J <- dnorm(p_grid, mean = 0.8, sd=0.2)
#prior_J <- rep( 1 , dens ) #flat

#Compute the likelihood at each value in grid
likelihood_J <- dbinom( 16000 , size=19800 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_J <- likelihood_J * prior_J

#Standardize the posterior (so it sums to 1)
posterior_J <- unstd.posterior_J / sum(unstd.posterior_J)

#Draw the plot
grid_data_J=data.frame(grid=p_grid,posterior=posterior_J,prior=prior_J,likelihood=likelihood_J)

ggplot(grid_data_J,aes(grid,posterior))+ geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
```

```{r}
#prior change in Mikkel
dens=100
p_grid <- seq( from=0 , to=1 , length.out=dens)

#Define the prior
prior_M <- dnorm(p_grid, mean= 0.8, sd=0.2)
#prior_M <- rep( 1 , dens ) #flat

#Compute the likelihood at each value in grid
likelihood_M <- dbinom( 6600 , size=13200 , prob=p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior_M <- likelihood_M * prior_M

#Standardize the posterior (so it sums to 1)
posterior_M <- unstd.posterior_M / sum(unstd.posterior_M)

#Draw the plot
grid_data_M=data.frame(grid=p_grid,posterior=posterior_M,prior=prior_M,likelihood=likelihood_M)

ggplot(grid_data_M,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")

```
Do you still see a difference between the results? Why?
Yes there is still a difference but it is very small. The reason for this is that the prior does not influence a large amount of data points. 


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
We would set the mean at change level (0.5) and have a large standard deviation. Our argument for this is that we assume they don't know anyhitng and therefore they will have 50% chance of answering correct. The variance has be to large because because we expect that when the level are at chance, the particpants may answer many questions wrong in a row.  


6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

```{r}

```

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.

```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Total=c(6,2,198,132),
  ID=c("Riccardo","Kristian","Josh","Mikkel"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



